---
title: "Regex"
output: html_document
---
# Part 1: Data Science Report
## Date: February 5, 2021
## GitHub: https://github.com/kaileemorrison/special-goggles

1. Import dataset 
```{r}
MData <- read.csv("./measurements.csv",header=TRUE) # loads dataset
```

2. Use a dplyr command and regular expressions to add a new column called Sp that shortens the genus name in the Organism column to only the first letter and period. (i.e. Homo sapies -> H. sapiens)
```{r}
library(dplyr) # loads dplyr package

X <- MData # define MData as new object 'X'

X$Sp <- sub("(\\w)\\w+","\\1.",MData$Organism) # creates new column that shortens genus name to first letter and period

X <- mutate(MData, Sp=sub("(\\w)\\w+","\\1.",MData$Organism)) # adds new column to the dataset

head(X) # checks that new 'Sp' column has been successfully added
```


3. Use a dplyr command to create a new dataset but with the Sp column instead of the organism column, moved to the beginning in place of the original Organism column. 
```{r}
X <- select(X, Sp, Limb.width,Limb.length,Observer) # selects the columns of interest and places 'Sp' at the beginning
head(X) # confirms selection and arrangement of columns
```


4. Save the dataset as a new file called MeasShort.csv
```{r}
write.csv(X,"MeasShort.csv") # saves dataset as a new file
```

